---
output:
  word_document: default
  html_document: default
---
---
title: "Dendrochronolgy in R"
author: "Jeff Atkins"
date: ' r format(Sys.time(), "%B %d, %Y)
output:
  word_document: default
  pdf_document: default
  html_document: default
  ---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Working with ring-width data

`dplR` version 1.7.2 is available on CRAN 

Here we show how to import tree-ring data.  

We are using data downloaded from the NOAA XXXX, specifically the Cook - Kelsey Tract TSCA-ITRDB NC005  [url - https://www.ncdc.noaa.gov/paleo-search/study/2987?siteId=15463]



### Installation of dplR

```{r install}
# install the package if not already installed and call it via library()
if(!require(dplR)){install.packages("dplR")}
library(dplR)
library(tidyverse)

```
## Importing ring-width data

The vignette included with dplR provides a thorough walkthrough (https://cran.r-project.org/web/packages/dplR/vignettes/intro-dplR.pdf) here we show some highlights based on the independet dataset we have downloaded.



```{r pressure, echo=FALSE}
# the primary function in dplR is read.rwl()
rw <- dplR::read.rwl("./data/dendro/nc005.rwl")

# lets look at the dimension of our dataset
dim(rw)   # 424 years and 31 series

# and you can see that the data are both class "rw" and "data.frame"
class(rw)

```

## Data descriptions

`dplR` has a function, `rw.report()` that provides a fairly detailed description of the data set(s) you are working with. 

```{r describe, echo=FALSE}
# make our report
rwl.report(rw)

```
## Plotting data (Spaghetti Plots)

We can now plot these data.

```{r plotting, echo=FALSE}

plot(rw, plot.type="spag")

```
Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.


## Detrending
```{r detrend}
## Detrending example, albeit brief.
rw.detrend <- detrend(rwl = rw, method = "ModNegExp", verbose = TRUE)

# we can look at the col names of the new object
names(rw.detrend)

# and we can look at the means of those indviduals
colMeans(rw.detrend, na.rm=TRUE)

```

## Building a Chronology

A chronology can be built from the detrended data. 

```{r chrono}
rw.crn <- chron(rw.detrend, prefix = "CAM")

# now we can plot these data
plot(rw.crn, add.spline=TRUE, nyrs=20)

```