---
output:
  word_document: default
  html_document: default
---
---
title: "FIA example in R"
author: "Jeff Atkins"
date: ' r format(Sys.time(), "%B %d, %Y)
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Installation of rFIA

```{r install}
# install the package if not already installed and call it via library()
if(!require(rFIA)){install.packages("rFIA")}
library(rFIA)
library(tidyverse)
library(sf)

```
## Getting started with rFIA

Please visit https://rfia.netlify.app/tutorial/downloading/ for a full run down on how to do this. Here, we work through their example using FIA data from the state of Virginia. 

You can also embed plots, for example:



```{r pressure, eval = FALSE}
## Download the state subset or Connecticut (requires an internet connection)
## Save as an object to automatically load the data into your current R session!
getFIA(states = 'VA', dir = './data/fia/')

## Get the forest type and forest type group reference tables
ref <- getFIA(states = 'ref', tables = c('FOREST_TYPE', 'FOREST_TYPE_GROUP'))

```


```{r read-FIA, echo=FALSE}
## But using the 'states' argument we can select individual states (or groups)
va <- readFIA(dir = './data/fia', states = 'VA')


# how many unique trees
va.tree <- tpa(va, grpBy = TREE)

va.tree <- tpa(va, byPlot = TRUE)

```
Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.



```{r data-time}


## Check spatial coverage of plots held in the database
plotFIA(va, plot.title = "FIA plot distribution, Virginia, USA" )

```



```{r clipping}
## let's make a bounding box
snp <- st_read("./data/fia/SNP_boundary/shen_PkBndryPly.shp")

## Most Recent Subset (2017)
snp.clip <- clipFIA(va, mask = snp) 


# let's look at some basic population estimates
tpaSNP <- tpa(snp.clip)
```


```{r by-plot}
# we can also analyze by plot
## Plot-level
snp_plot <- tpa(snp.clip, byPlot = TRUE)

## or by tree
## Plot-level
snp_plot_tree <- tpa(snp.clip, byPlot = TRUE, grpBy = TREE)

## Size class
## Group estimates by size class
snp_sizeClass <- tpa(snp.clip, bySizeClass = TRUE)

## Using our estimates from above (all inventory years in RI)
plotFIA(snp_sizeClass, y = BAA, x = sizeClass, plot.title = 'Size class distribution of trees in SNP')
```